### 2017-12-10 linearity-2018/adjoint-models.tex
More writing.
I was thinking that we could also include associtivity here, but it is
proving more difficult than I thought.

We first introduce adjoint models, and the two particular adjoint
models that will be used to model the Lambek Calculus with and without
associativity.  However, before we can do this we have to define a
very simple categorical structure known as a magmoidal category with a
unit.
%% \begin{definition}
%%   \label{def:magmoidal-categories}
%%   1
%% \end{definition}

>>>>

We first introduce adjoint models, and the two particular adjoint
models that will be used to model the Lambek Calculus with and without
associativity.  However, before we can do this we have to define a
very simple categorical structure known as a magmoidal category with a
unit.
\begin{definition}
  \label{def:magmoidal-categories}
  A \textbf{magmoidal category with a unit (MCU)}, $(\cat{M}, \otimes,
  I, u_l,u_r)$, is a category $\cat{M}$ with a distinguished object,
  $I$, called the unit, a bifunctor $\otimes : \cat{M} \times \cat{M}
  \to \cat{M}$, and two natural isomorphisms $u_l : I \otimes A \mto
  A$ and $u_r : A \otimes I \mto A$ called the left and right unitors
  respectively.
\end{definition}
\noindent
A MCU is then a generalization of a monoidal category by removing the
associator.  This is necessary in order to model the non-associative
Lambek calculus.

### 2017-12-10 linearity-2018/introduction.tex
More writing.
I was thinking that we could also include associtivity here, but it is
proving more difficult than I thought.

Linguists have wondered if a modality for exchange can be added to the
Lambek Calculus similarly to how the of-course! modality adds
weakening and contraction to linear logic.  de Paiva and
Eades~\cite{?} show that this is possible by adding a modality $\kappa
A$ with the property that $(\kappa A \otimes B) \multimap (B \otimes
\kappa A)$.  The adjoint model introduced above induces the modality
$eA = F(G\,A)$ with the property $eA \otimes eB \multimap eB \otimes
eA$.  Thus providing a second solution to the problem, but we go
further, and show such a modality can be given for associativity. It
arises as the induced comonad of the adjoint model where $\cat{M}_1$
is a monoidal category, and $\cat{M}_2$ is a vanilla magmoidal
category with a unit.  Then we will obtain the isomorphism $(aA
\otimes aB) \otimes aC) \multimap aA \otimes (aB \otimes aC)$.

@@

We introduce the idea above of having a modality for exchange and one
for associativity, but what about weakening and contraction?  Indeed
it is possible to give modalities for just weakening, $wA$, and just
contraction, $cA$, using adjoint models.  Now that we have each
structural rule isolated into their own modality is it possible to put
them together to form new modalities that combine structural rules?
The answer to this question has already been show to be positive, at
least for weakening and contraction, but we extend this line of work
to include exchange and associativity as well.
Jacobs~\cite{JACOBS199473} used monads and distributive laws to relate
modalities, but Melli{\'e}s~\cite{Mellies:2004} shows how to use
adjunctions to accomplish the same, but in a more intuitive and
natural way.  Thus, this part of our work is a natural extension of
Melli{\'e}s'.  At this point we are starting to see that this is a
very general framework.

>>>>

Linguists have wondered if a modality for exchange can be added to the
Lambek Calculus similarly to how the of-course! modality adds
weakening and contraction to linear logic.  de Paiva and
Eades~\cite{?} show that this is possible by adding a modality $\kappa
A$ with the property that $(\kappa A \otimes B) \multimap (B \otimes
\kappa A)$.  The adjoint model introduced above induces the modality
$eA = F(G\,A)$ with the property $eA \otimes eB \multimap eB \otimes
eA$.  Thus providing a second solution to the problem.

@@

We introduce the idea above of having a modality for exchange, but
what about associativity, weakening and contraction?  Indeed it is
possible to give modalities for these structural rules as well using
adjoint models.  Now that we have each structural rule isolated into
their own modality is it possible to put them together to form new
modalities that combine structural rules?  The answer to this question
has already been show to be positive, at least for weakening and
contraction, but we extend this line of work to include exchange, and
in the future associativity as well.  Jacobs~\cite{JACOBS199473} used
monads and distributive laws to relate modalities, but
Melli{\'e}s~\cite{Mellies:2004} shows how to use adjunctions to
accomplish the same, but in a more intuitive and natural way.  Thus,
this part of our work is a natural extension of Melli{\'e}s'.  At this
point we are starting to see that this is a very general framework.
